---
title: "Gender Experiment - Power Test - w241 Final Project"
author: "Daniel Alvarez, Bethany Keller, Austin Doolittle"
date: "3/30/2020"
output:
    github_document: default
    pdf_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(
    inputFile, encoding = encoding,
    output_format = c('github_document', 'pdf_document'))
    })
# output:
#   rmarkdown::github_document:
#       toc: true
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r,include=FALSE, results='hide'}
# clear out environment variables
rm(list=ls())

# set seed for randomization
set.seed(1)
```


Load packages
```{r, include=FALSE}
# load packages 
library(foreign)
library(data.table)
library(knitr)
library(dplyr)
library(ggplot2)
library(foreign)
library(zoo)
library(lmtest) # robust standard errors
library(sandwich) # robust standard errors
library(stargazer) # reporting formatted results
library(magrittr) # printing document
library(pwr) # estimating power
```

Write repeatable functions used throughout this exercise.

```{r write repeatable functions}
# show mean comply rates by assignment status in the pilot
complyratesbytreatment <- function(dt) { 
  return(dt[,.('mean_comply_rate'=mean(comply)),keyby=.(assignment_status)])
  }

# estimate treatment effect for the male voice treatment
ate_treatmentmale <- function(dt) {
  return(mean(dt[dt$assignment_status=='TM',]$comply)-mean(dt[dt$assignment_status=='C',]$comply))
}

# estimate treatment effect for the female voice treatment
ate_treatmentfemale <- function(dt) {
  return(mean(dt[dt$assignment_status=='TF',]$comply)-mean(dt[dt$assignment_status=='C',]$comply))
}

# estimate effect between male and female voice treatments
ate_treatment_btwnmalefemale <- function(dt) {
  return(mean(dt[dt$assignment_status=='TM',]$comply)-mean(dt[dt$assignment_status=='TF',]$comply))
}

# conduct a two-sample t-test between male audio treatment and control
t_test_result <- function(dt, treatment, control) {
  # vectors with bids for control and treatment groups
  optimal_t <- dt[dt$assignment_status==treatment, comply]
  optimal_c <- dt[dt$assignment_status==control, comply]
  return(t.test(optimal_t, optimal_c))
}

# regression of the comply rate on the assignment status with covariate adjustment for subject gender and age
reg <- function(data){
  data[,lm(comply_rate ~ as.factor(assignment_status)+gender+age)]
}

# regression of the comply rate on the assignment status with covariate adjustment for subject gender and age and interaction terms
reg_interaction <- function(data){
  data[,lm(comply_rate ~ as.factor(assignment_status)+gender+age+gender*as.factor(assignment_status)+age*as.factor(assignment_status))]
}

# report regression results with robust standard errors
reg_agg_robust <- function(data){
  reg <- data[,lm(comply_rate ~ as.factor(assignment_status)+gender+age)]
  return(coeftest(reg, vcovHC(reg, type='HC3')))
}

# report regression results with robust standard errors and interaction terms
reg_agg_robust_interaction <- function(data){
  reg <- data[,lm(comply_rate ~ as.factor(assignment_status)+gender+age+gender*as.factor(assignment_status)+age*as.factor(assignment_status))]
  return(coeftest(reg, vcovHC(reg, type='HC3')))
}

# regression of the comply rate on the assignment status with covariate adjustment for subject gender and age
# report regression results with clustered standard errors at subject ID level
reg_agg_clustered <- function(data){
  reg <- data[,lm(comply_rate ~ as.factor(assignment_status)+gender+age)]
  return(coeftest(reg, vcovCL(reg, cluster = data[ ,subject_id])))
}

# report regression results with clustered standard errors at subject ID level and interaction terms
reg_agg_clustered_interaction <- function(data){
  reg <- data[,lm(comply_rate ~ as.factor(assignment_status)+gender+age+gender*as.factor(assignment_status)+age*as.factor(assignment_status))]
  return(coeftest(reg, vcovCL(reg, cluster = data[ ,subject_id])))
}

# logistic regression of the comply rate on the assignment status with covariate adjustment for subject gender and age
logistic_reg <- function(data){
  data[,glm(comply ~ as.factor(assignment_status)+gender+age+as.factor(subject_id), family = "binomial")]
} 

# report logistic regression results with robust standard errors
logistic_reg_robust <- function(data){
  logistic <- data[,glm(comply ~ as.factor(assignment_status)+gender+age+as.factor(subject_id), family = "binomial")]
  return(coeftest(logistic,vcovHC(logistic, type='HC3')))
}

# report logistic regression results with clustered standard errors
logistic_reg_clustered <- function(data){
  logistic <- data[,glm(comply ~ as.factor(assignment_status)+gender+age+as.factor(subject_id), family = "binomial")]
  return(coeftest(logistic,vcovHC(logistic, cluster = data[ ,subject_id])))
}

# robust standard errors
robust_se <- function(mod, type = 'HC3') { 
  sqrt(diag(vcovHC(mod, type)))
}

# clustered standard errors
cluster_se <- function(mod, data) { 
  sqrt(diag(vcovCL(mod, cluster=data[,subject_id])))
}

```

## Analyze all moves data

We performed several analyses on the all moves dataset comprising each suggested move outcome, as follows:
- 1. Analysis of comply rates for optimal suggested moves
- 2. Analysis of comply rates for non-optimal suggested moves
- 3. Analysis of comply rates by game number (in order to examine whether there is a trend in compliance rates as the number of games played increases)
- 4. Analysis of comply rates for order of play moves (in order to examine whether compliance increases with play move number)
- 5. Analysis of comply rates by gender of subject


Read in all moves.

```{r}
all_moves <- fread('./all_moves_final.csv')
all_moves
```

Create binary comply variable for if suggested move was taken or not
```{r}
# create a binary comply variable for if suggested move was taken or not
all_moves[, comply:=ifelse((suggested_move_row==move_taken_row)&(suggested_move_column==move_taken_column),1,0)] 
```


Explore data
```{r}
str(all_moves)

summary(all_moves)
```

Create subsets of the data for optimal and non-optimal suggested moves

```{r}
# filter out compliers in the study and optimal suggested moves
all_moves_optimal <- all_moves[(experiment_status == 'C') & (is_suggested_move_optimal==TRUE)]
all_moves_nonoptimal <- all_moves[(experiment_status == 'C') & (is_suggested_move_optimal==FALSE)]

# number of rows
nrow(all_moves_optimal)
nrow(all_moves_nonoptimal)
```


Create an aggregate dataset with mean optimal suggested move rates per subject

```{r}
# collapse the dataset by mean optimal move
all_moves_agg <- all_moves[experiment_status == 'C', list(mean_optimal_move=mean(is_suggested_move_optimal, na.rm=T)), keyby=subject_id]
head(all_moves_agg)
nrow(all_moves_agg)

```

Create an aggregate dataset with mean compliance rates per subject for optimal suggested moves data subset.

```{r}
# collapse the dataset by mean optimal move
optimal_moves_agg <- all_moves_optimal[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T), assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

# show snapshot of aggregated dataset
head(optimal_moves_agg)
nrow(optimal_moves_agg)

# show distribution of compliance rates by subject
hist(optimal_moves_agg$comply_rate, main="Histogram of compliance rates by subject for optimal moves", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=50)
```

Create an aggregate dataset with mean compliance rates per subject for non-optimal suggested moves data subset.

```{r}
# collapse the dataset by mean optimal move
nonoptimal_moves_agg <- all_moves_nonoptimal[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T), assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

# show snapshot of aggregated dataset
head(nonoptimal_moves_agg)
nrow(nonoptimal_moves_agg)

# show distribution of non-compliance rates by subject
hist(nonoptimal_moves_agg$comply_rate, main="Histogram of compliance rates by subject for non-optimal moves", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=50)

```



## 1. Analysis of comply rates for optimal suggested moves

See distribution of optimal moves for each subject. Some subject received more optimal moves than others, but distribution of proportion of optimal suggested moves by subject should be broadly symmetric since move suggestions were random.

```{r}
# cross tab subject id and boolean value for optimal move suggestion
table(all_moves$subject_id, all_moves$is_suggested_move_optimal)

# show distribution of proportion of optimal moves by subject
hist(all_moves_agg$mean_optimal_move, main="Histogram of proportion of suggested optimal moves", xlab="Proportion of suggested optimal moves", ylab = "Frequency", breaks=50)
```

### Estimate treatment effects for optimal move and non-optimal move datasets separately

#### Estimate the treatment effect among the optimal move suggestion recipients



```{r estimate comply rates}

complyratesbytreatment(dt=all_moves_optimal)
ate_treatmentmale(dt=all_moves_optimal)
ate_treatmentfemale(dt=all_moves_optimal)
ate_treatment_btwnmalefemale(dt=all_moves_optimal)

# report average treatment effects output in nicer format
kable(complyratesbytreatment(dt=all_moves_optimal))
```


For the optimal suggested moves only, conduct a t-test to observe differences between the treatments (male voice and female voices) mean and the control mean of proportions.


```{r two-sample t-test}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_optimal_tm <- t_test_result(all_moves_optimal, 'TM', 'C')
t_test_result_optimal_tf <- t_test_result(all_moves_optimal, 'TF', 'C')
t_test_result_optimal_t <- t_test_result(all_moves_optimal, 'TM', 'TF')

t_test_result_optimal_tm
t_test_result_optimal_tf
t_test_result_optimal_t
```

We observe from the outcome data for optimal suggested moves that the treatment effect for the male voice treatment is statistically significant with a p-value in the t-test of `r t_test_result_optimal_tm$p.value`. Also, the treatment effect for female voice treatment is statistically different from zero with a p-value in the t-test of `r t_test_result_optimal_tf$p.value`. The t-test comparing the proportions for the male and female voice treatment is not statistically different from zero.

The linear regression of the aggregated comply rate on assignment status adjusting for the subject gender and age covariates and with subject id-level fixed effects tells gives back the estimated treatment effect.

```{r linear regression for optimal moves}
# linear regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with robust standard errors
reg_agg_robust(data=optimal_moves_agg)

# linear regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with clustered standard errors
reg_agg_clustered(data=optimal_moves_agg)

```

Report the formatted linear regression results on aggregated data for optimal move data.

```{r}
# report the formatted regression results
stargazer(reg(data=optimal_moves_agg), reg(data=optimal_moves_agg), reg_interaction(data=optimal_moves_agg),reg_interaction(data=optimal_moves_agg),
          type = 'text', 
          se = list(robust_se(reg(data=optimal_moves_agg)), cluster_se(reg(data=optimal_moves_agg),optimal_moves_agg),robust_se(reg_interaction(data=optimal_moves_agg)), cluster_se(reg_interaction(data=optimal_moves_agg),optimal_moves_agg)),
          add.lines = list(c('SE', 'Robust', 'Clustered', 'Robust', 'Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)
```



The logistic regression of the binary comply outcome on assignment status adjusting for the subject gender and age covariates and with subject id-level fixed effects might tell a more nuanced story. Report results with both robust standard errors and clustering at the subject ID level. Clustering at the subject ID level is important since the number of the moves (and therefore, number of observations) may differ across subjects. Clustering ensures that the standard errors reflects the tighter variation introduced at the subject identifier level.

```{r logistic regression for optimal moves}
# logistic regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with robust standard errors
logistic_reg_robust(data=all_moves_optimal)


# logistic regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with clustered standard errors on subject id
logistic_reg_clustered(data=all_moves_optimal)

```

Report the formatted logistic regression results for optimal move data.

```{r}
# report the formatted regression results
stargazer(logistic_reg(data=all_moves_optimal), logistic_reg(data=all_moves_optimal),
          type = 'text', 
          se = list(robust_se(logistic_reg(data=all_moves_optimal)), cluster_se(logistic_reg(data=all_moves_optimal),all_moves_optimal)),
          add.lines = list(c('SE', 'Robust', 'Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)

```


Convert coefficients of logistic regression into odds ratios.

```{r}
# convert the coefficients into odds ratios
exp(coef(logistic_reg(data=all_moves_optimal)))

## odds ratios and 95% CI
kable(exp(cbind(OR = coef(logistic_reg(data=all_moves_optimal)), confint(logistic_reg(data=all_moves_optimal)))))
```


## 2. Analysis of comply rates for non-optimal suggested moves

#### Estimate the treatment effect among the non-optimal move suggestion recipients

```{r estimate comply rates}
complyratesbytreatment(dt=all_moves_nonoptimal)
ate_treatmentmale(dt=all_moves_nonoptimal)
ate_treatmentfemale(dt=all_moves_nonoptimal)
ate_treatment_btwnmalefemale(dt=all_moves_nonoptimal)

# report average treatment effects output in nicer format
kable(complyratesbytreatment(dt=all_moves_nonoptimal))
```

For the optimal suggested moves only, conduct a t-test to observe differences between the treatments (male voice and female voices) mean and the control mean of proportions.


```{r two-sample t-test}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_nonoptimal_tm <- t_test_result(all_moves_nonoptimal, 'TM', 'C')
t_test_result_nonoptimal_tf <- t_test_result(all_moves_nonoptimal, 'TF', 'C')
t_test_result_nonoptimal_t <- t_test_result(all_moves_nonoptimal, 'TM', 'TF')

t_test_result_nonoptimal_tm
t_test_result_nonoptimal_tf
t_test_result_nonoptimal_t 
```

We observe from the pilot for non-optimal suggested moves that there the treatment effect for the male voice treatment is not statistically different from zero with a p-value in the t-test of `r t_test_result_nonoptimal_tm$p.value`. However, the treatment effect for female voice treatment is statistically different from zero with a p-value in the t-test of `r t_test_result_nonoptimal_tf$p.value`.

The linear regression of the aggregated comply rate on assignment status adjusting for the subject gender and age covariates and with subject id-level fixed effects tells gives back the estimated treatment effect.

```{r linear regression for non-optimal moves}
# linear regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with robust standard errors
reg_agg_robust(data=nonoptimal_moves_agg)

# linear regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with clustered standard errors
reg_agg_clustered(data=nonoptimal_moves_agg)

```

Report the formatted linear regression results on aggregated data for nonoptimal move data.

```{r}
# report the formatted regression results
stargazer(reg(data=nonoptimal_moves_agg), reg(data=nonoptimal_moves_agg), reg_interaction(data=nonoptimal_moves_agg), reg_interaction(data=nonoptimal_moves_agg),
          type = 'text', 
          se = list(robust_se(reg(data=nonoptimal_moves_agg)), cluster_se(reg(data=nonoptimal_moves_agg),nonoptimal_moves_agg),robust_se(reg_interaction(data=nonoptimal_moves_agg)), cluster_se(reg_interaction(data=nonoptimal_moves_agg),nonoptimal_moves_agg)),
          add.lines = list(c('Move Type', 'Nonoptimal Moves', 'Nonoptimal Moves', 'Nonoptimal Moves', 'Nonoptimal Moves'),c('SE', 'Robust', 'Clustered', 'Robust', 'Clustered')),
          omit.stat = c('ser', 'F'),
          header=F)
```

The logistic regression of the binary comply outcome on assignment status adjusting for the subject gender and age covariates and with subject id-level fixed effects might tell a more nuanced story. Report results with both robust standard errors and clustering at the subject ID level.

```{r logistic regression for non-optimal moves}
# logistic regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with robust standard errors
logistic_reg_robust(data=all_moves_nonoptimal)


# logistic regression of the comply rate on the assignment status with covariate adjustment for subject gender and age reported with clustered standard errors on subject id
logistic_reg_clustered(data=all_moves_nonoptimal)

```

Report the formatted regression results for logistical regression for logistic regression.

```{r formatted logistic regression results for non-optimal moves}
# report the formatted regression results
stargazer(logistic_reg(data=all_moves_nonoptimal), logistic_reg(data=all_moves_nonoptimal),
          type = 'text', 
          se = list(robust_se(logistic_reg(data=all_moves_nonoptimal)), cluster_se(logistic_reg(data=all_moves_nonoptimal),all_moves_nonoptimal)),
          add.lines = list(c('SE', 'Robust', 'Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)
```

Convert coefficients of logistic regression into odds ratios.

```{r}
# convert the coefficients into odds ratios
exp(coef(logistic_reg(data=all_moves_nonoptimal)))

## odds ratios and 95% CI
#kable(exp(cbind(OR = coef(logistic_reg(data=all_moves_nonoptimal)), confint(logistic_reg(data=all_moves_nonoptimal)))))
```

## 3. Analysis of comply rates by game number 

This analysis is conducted in order to examine whether there is a trend in compliance rates as the number of games played increases. We might suspect that compliance decreases with continued game play differentially by type of treatment.

Conduct diagnostic checks on the game number data. Note that there is 1 subject that attritted by not completing the fifth game. This was just observed as due to a technical bug issue in the pilot game play whereby a player could opt out of the final game and go right to the post-treatment survey evaluation. This was subsequently modified changed in the full experimental roll-out. 

```{r diagnostic checks on game number data}
# check unique number of game numbers
kable(unique(all_moves$game_number))

# check number of moves played for each game number
kable(table(all_moves$game_number))

# cross-tab number of moves played for each game number by treatment assignment
kable(table(all_moves$game_number, all_moves$assignment_status))

# assess attrition if not all players completed each game
kable(table(all_moves$game_number, all_moves$subject_id))
```

First, create subsets of the data corresponding to each game.


```{r subset data corresponding to game number, include=FALSE}
all_moves_game0 <- all_moves[game_number == 0]
all_moves_game1 <- all_moves[game_number == 1]
all_moves_game2 <- all_moves[game_number == 2]
all_moves_game3 <- all_moves[game_number == 3]
all_moves_game4 <- all_moves[game_number == 4]
```


Create aggregated datasets for each of the `all moves subsets` by subject ID.

```{r create aggregated datasets}
# collapse the datasets by mean compliance by subject ID
all_moves_game0_agg <- all_moves_game0[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_game1_agg <- all_moves_game1[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_game2_agg <- all_moves_game2[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_game3_agg <- all_moves_game3[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_game4_agg <- all_moves_game4[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

# show snapshot of aggregated dataset
nrow(all_moves_game0_agg)
nrow(all_moves_game1_agg)
nrow(all_moves_game2_agg)
nrow(all_moves_game3_agg)
nrow(all_moves_game4_agg)
```

Examine compliance distributions by game.

```{r histogram compliance rates for each game by subject}
# show distribution of compliance rates for each game by subject
par(mfrow=c(3,2))

hist(all_moves_game0_agg$comply_rate, main="Compliance rates by subject for Game 1", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=20)

hist(all_moves_game1_agg$comply_rate, main="Compliance rates by subject for Game 2", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=20)

hist(all_moves_game2_agg$comply_rate, main="Compliance rates by subject for Game 3", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=20)

hist(all_moves_game3_agg$comply_rate, main="Compliance rates by subject for Game 4", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=20)

hist(all_moves_game4_agg$comply_rate, main="Compliance rates by subject for Game 5", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=20)
```



#### Estimate the treatment effects for each game number

```{r treatment effects for first game}
complyratesbytreatment(dt=all_moves_game0)
ate_treatmentmale(dt=all_moves_game0)
ate_treatmentfemale(dt=all_moves_game0)
ate_treatment_btwnmalefemale(dt=all_moves_game0)

# report mean comply rates in nice format
kable(complyratesbytreatment(dt=all_moves_game0))
```


```{r treatment effects for second game}
complyratesbytreatment(dt=all_moves_game1)
ate_treatmentmale(dt=all_moves_game1)
ate_treatmentfemale(dt=all_moves_game1)
ate_treatment_btwnmalefemale(dt=all_moves_game1)

# report mean comply rates in nice format
kable(complyratesbytreatment(dt=all_moves_game1))

```


```{r treatment effects for third game}
complyratesbytreatment(dt=all_moves_game2)
ate_treatmentmale(dt=all_moves_game2)
ate_treatmentfemale(dt=all_moves_game2)
ate_treatment_btwnmalefemale(dt=all_moves_game2)

# report mean comply rates in nice format
kable(complyratesbytreatment(dt=all_moves_game2))
```


```{r treatment effects for fourth game}
complyratesbytreatment(dt=all_moves_game3)
ate_treatmentmale(dt=all_moves_game3)
ate_treatmentfemale(dt=all_moves_game3)
ate_treatment_btwnmalefemale(dt=all_moves_game3)

# report mean comply rates in nice format
kable(complyratesbytreatment(dt=all_moves_game3))
```


```{r treatment effects for fifth game}
complyratesbytreatment(dt=all_moves_game4)
ate_treatmentmale(dt=all_moves_game4)
ate_treatmentfemale(dt=all_moves_game4)
ate_treatment_btwnmalefemale(dt=all_moves_game4)

# report mean comply rates in nice format
kable(complyratesbytreatment(dt=all_moves_game4))

```

Conduct t-test by game
```{r two-sample t-test for first game}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_game0_tm <- t_test_result(all_moves_game0, 'TM', 'C')
t_test_result_game0_tf <- t_test_result(all_moves_game0, 'TF', 'C')
t_test_result_game0_t <- t_test_result(all_moves_game0, 'TM', 'TF')

t_test_result_game0_tm
t_test_result_game0_tf
t_test_result_game0_t
```

```{r two-sample t-test for second game}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_game1_tm <- t_test_result(all_moves_game1, 'TM', 'C')
t_test_result_game1_tf <- t_test_result(all_moves_game1, 'TF', 'C')
t_test_result_game1_t <- t_test_result(all_moves_game1, 'TM', 'TF')

t_test_result_game1_tm
t_test_result_game1_tf
t_test_result_game1_t
```

```{r two-sample t-test for third game}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_game2_tm <- t_test_result(all_moves_game2, 'TM', 'C')
t_test_result_game2_tf <- t_test_result(all_moves_game2, 'TF', 'C')
t_test_result_game2_t <- t_test_result(all_moves_game2, 'TM', 'TF')

t_test_result_game2_tm
t_test_result_game2_tf
t_test_result_game2_t

```


```{r two-sample t-test for fourth game}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_game3_tm <- t_test_result(all_moves_game3, 'TM', 'C')
t_test_result_game3_tf <- t_test_result(all_moves_game3, 'TF', 'C')
t_test_result_game3_t <- t_test_result(all_moves_game3, 'TM', 'TF')

t_test_result_game3_tm
t_test_result_game3_tf
t_test_result_game3_t
```


```{r two-sample t-test for fifth game}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_game4_tm <- t_test_result(all_moves_game4, 'TM', 'C')
t_test_result_game4_tf <- t_test_result(all_moves_game4, 'TF', 'C')
t_test_result_game4_t <- t_test_result(all_moves_game4, 'TM', 'TF')

t_test_result_game4_tm
t_test_result_game4_tf
t_test_result_game4_t
```

Show results of linear regressions run on aggregated data for all 5 games with clustered standard errors.

```{r}
# report the formatted regression results for all games with clustered standard errors.
stargazer(reg(all_moves_game0_agg), reg(all_moves_game1_agg),reg(all_moves_game2_agg),reg(all_moves_game3_agg),reg(all_moves_game4_agg),
          type = 'text', 
          se = list(cluster_se(reg(data=all_moves_game0_agg),all_moves_game0_agg),cluster_se(reg(data=all_moves_game1_agg),all_moves_game1_agg),cluster_se(reg(data=all_moves_game2_agg),all_moves_game2_agg),cluster_se(reg(data=all_moves_game3_agg),all_moves_game3_agg),cluster_se(reg(data=all_moves_game4_agg),all_moves_game4_agg)),
          add.lines = list(c('Games','Game1','Game2','Game3','Game4','Game5'),c('SE', 'Clustered', 'Clustered','Clustered','Clustered','Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)


```




Show results of logistic regressions run on by move data for all 5 games with clustered standard errors (clustered by subject id).

```{r}
# report the formatted regression results
stargazer(logistic_reg(all_moves_game0), logistic_reg(all_moves_game1), logistic_reg(all_moves_game2),logistic_reg(all_moves_game3),logistic_reg(all_moves_game4),
          type = 'text', 
          se = list(cluster_se(logistic_reg(all_moves_game0),all_moves_game0), cluster_se(logistic_reg(all_moves_game1),all_moves_game1),cluster_se(logistic_reg(all_moves_game2),all_moves_game2), cluster_se(logistic_reg(all_moves_game3),all_moves_game3),cluster_se(logistic_reg(all_moves_game4),all_moves_game4)), add.lines = list(c('Game No.','Game 1','Game 2','Game 3','Game 4','Game 5'),c('SE', 'Clustered', 'Clustered','Clustered','Clustered','Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)


```

# convert the coefficients of logistic regression into odds ratios

```{r convert coefficients to odd ratios}
kable(cbind(exp(coef(logistic_reg(all_moves_game0))),exp(coef(logistic_reg(all_moves_game1))),exp(coef(logistic_reg(all_moves_game2))),exp(coef(logistic_reg(all_moves_game3))),exp(coef(logistic_reg(all_moves_game4)))))

```


## 4. Analysis of comply rates for order of play moves

The analysis is conducted in order to examine whether compliance increases with play move number.


Create dataset for first moves only

```{r create datasets for first, second and third moves only}
all_moves_move1 <- all_moves[(game_number == 0 & move_number ==1) | (game_number == 1 & move_number ==2) |
                             (game_number == 2 & move_number ==1) | (game_number == 3 & move_number ==2) |
                             (game_number == 4 & move_number ==1)]

all_moves_move2 <- all_moves[(game_number == 0 & move_number ==3) | (game_number == 1 & move_number ==4) |
                             (game_number == 2 & move_number ==3) | (game_number == 3 & move_number ==4) |
                             (game_number == 4 & move_number ==3)]

all_moves_move3 <-all_moves[(game_number == 0 & move_number ==5) | (game_number == 1 & move_number ==6) |
                             (game_number == 2 & move_number ==5) | (game_number == 3 & move_number ==6) |
                             (game_number == 4 & move_number ==5)]
```

Created aggregated datasets from the ordered move datasets.

```{r create aggregated datasets}
# collapse the datasets by mean compliance by subject ID
all_moves_move1_agg <- all_moves_move1[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_move2_agg <- all_moves_move2[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_move3_agg <- all_moves_move3[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

```

Examine compliance distributions by ordered move.

```{r histogram compliance rates for each game by subject}
# show distribution of compliance rates for each game by subject
par(mfrow=c(3,1))

hist(all_moves_move1_agg$comply_rate, main="Compliance rates by subject for Move 1", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=30)

hist(all_moves_move2_agg$comply_rate, main="Compliance rates by subject for Move 2", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=30)

hist(all_moves_move3_agg$comply_rate, main="Compliance rates by subject for Move 3", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=30)

```


Run diagnostic checks.

```{r diagnostics on datasets}
# check number of moves played for each game number (should be one)
kable(table(all_moves_move1$game_number))
kable(table(all_moves_move2$game_number))
kable(table(all_moves_move3$game_number))

# cross-tab number of moves played for each game number by treatment assignment
kable(table(all_moves_move1$game_number, all_moves_move1$assignment_status))
kable(table(all_moves_move2$game_number, all_moves_move2$assignment_status))
kable(table(all_moves_move3$game_number, all_moves_move3$assignment_status))

# assess attrition if not all players made a first, second or third move
kable(table(all_moves_move1$game_number, all_moves_move1$subject_id))
kable(table(all_moves_move2$game_number, all_moves_move2$subject_id))
kable(table(all_moves_move3$game_number, all_moves_move3$subject_id))
```

Estimate treatment effects

```{r estimate treatment effect for first moves only}
complyratesbytreatment(dt=all_moves_move1)
ate_treatmentmale(dt=all_moves_move1)
ate_treatmentfemale(dt=all_moves_move1)
ate_treatment_btwnmalefemale(dt=all_moves_move1)

kable(complyratesbytreatment(dt=all_moves_move1))
```


```{r estimate treatment effect for second moves only}
complyratesbytreatment(dt=all_moves_move2)
ate_treatmentmale(dt=all_moves_move2)
ate_treatmentfemale(dt=all_moves_move2)
ate_treatment_btwnmalefemale(dt=all_moves_move2)

kable(complyratesbytreatment(dt=all_moves_move2))
```


```{r estimate treatment effect for third moves only}
complyratesbytreatment(dt=all_moves_move3)
ate_treatmentmale(dt=all_moves_move3)
ate_treatmentfemale(dt=all_moves_move3)
ate_treatment_btwnmalefemale(dt=all_moves_move3)

kable(complyratesbytreatment(dt=all_moves_move3))
```

Conduct t-tests by move

```{r two-sample t-test for first moves only}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_move1_tm <- t_test_result(all_moves_move1, 'TM', 'C')
t_test_result_move1_tf <- t_test_result(all_moves_move1, 'TF', 'C')
t_test_result_move1_t <- t_test_result(all_moves_move1, 'TM', 'TF')

t_test_result_move1_tm
t_test_result_move1_tf
t_test_result_move1_t
```


```{r two-sample t-test for second moves only}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_move2_tm <- t_test_result(all_moves_move2, 'TM', 'C')
t_test_result_move2_tf <- t_test_result(all_moves_move2, 'TF', 'C')
t_test_result_move2_t <- t_test_result(all_moves_move2, 'TM', 'TF')

t_test_result_move2_tm
t_test_result_move2_tf
t_test_result_move2_t
```


```{r two-sample t-test for second moves only}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_move3_tm <- t_test_result(all_moves_move3, 'TM', 'C')
t_test_result_move3_tf <- t_test_result(all_moves_move3, 'TF', 'C')
t_test_result_move3_t <- t_test_result(all_moves_move3, 'TM', 'TF')

t_test_result_move3_tm
t_test_result_move3_tf
t_test_result_move3_t
```

Show results of linear regressions run on aggregated data for all 3 moves with robust standard errors.

```{r}
# report the formatted regression results
stargazer(reg(all_moves_move1_agg), reg(all_moves_move2_agg),reg(all_moves_move3_agg),
          type = 'text', 
          se = list(robust_se(reg(all_moves_move1_agg)),robust_se(reg(all_moves_move2_agg)),robust_se(reg(all_moves_move3_agg))),
          add.lines = list(c('Game No.','Move 1','Move 2','Move 3'),c('SE', 'Robust', 'Robust','Robust')),
          #omit.stat = c('ser', 'F'),
          header=F)


```

Show results of linear regressions run on aggregated data for all 3 moves with clustered standard errors.

```{r}
# report the formatted regression results
stargazer(reg(all_moves_move1_agg), reg(all_moves_move2_agg),reg(all_moves_move3_agg),
          type = 'text', 
          se = list(cluster_se(reg(all_moves_move1_agg),all_moves_move1_agg),cluster_se(reg(all_moves_move2_agg),all_moves_move2_agg),cluster_se(reg(all_moves_move3_agg),all_moves_move3_agg)),
          add.lines = list(c('Move Order No.','Move 1','Move 2','Move 3'),c('SE', 'Clustered', 'Clustered','Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)


```


Show results of logistic regressions run on by move data for the three ordered moves datasets with clustered standard errors (clustered by subject id).

```{r}
# report the formatted regression results
stargazer(logistic_reg(all_moves_move1), logistic_reg(all_moves_move2), logistic_reg(all_moves_move3),
          type = 'text', 
          se = list(cluster_se(logistic_reg(all_moves_move1),all_moves_move1), cluster_se(logistic_reg(all_moves_move2),all_moves_move2),cluster_se(logistic_reg(all_moves_move3),all_moves_move3)), add.lines = list(c('Game No.','Move 1','Move 2','Move 3'),c('SE', 'Clustered', 'Clustered','Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)


```


## 4. Analysis of comply rates by gender of subject participant in the study

The analysis is conducted in order to examine whether treatment effects differ by subject gender


Create datasets by subject gender 

```{r create datasets for first, second and third moves only}
all_moves_male <- all_moves[gender=='M']

all_moves_female <- all_moves[gender=='F']

#count number of observations
nrow(all_moves_male)
nrow(all_moves_female)
```

Created aggregated datasets from the by-gender datasets.

```{r create aggregated datasets}
# collapse the datasets by mean compliance by subject ID
all_moves_male_agg <- all_moves_male[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

all_moves_female_agg <- all_moves_female[experiment_status == 'C', list(comply_rate=mean(comply, na.rm=T),assignment_status=unique(assignment_status), gender=unique(gender), age=unique(age)), keyby=subject_id]

```

Examine compliance distributions by-gender

```{r histogram compliance rates for each game by subject}
# show distribution of compliance rates for each subject gender
par(mfrow=c(2,1))

hist(all_moves_male_agg$comply_rate, main="Compliance rates for male subjects", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=30)

hist(all_moves_female_agg$comply_rate, main="Compliance rates for female subjects", xlab="Proportion of compliance rates", ylab = "Frequency", breaks=30)

```

Estimate treatment effects

```{r estimate treatment effect for male subjects}
complyratesbytreatment(dt=all_moves_male)
ate_treatmentmale(dt=all_moves_male)
ate_treatmentfemale(dt=all_moves_male)
ate_treatment_btwnmalefemale(dt=all_moves_male)

kable(complyratesbytreatment(dt=all_moves_male))
```


```{r estimate treatment effect for female subjects}
complyratesbytreatment(dt=all_moves_female)
ate_treatmentmale(dt=all_moves_female)
ate_treatmentfemale(dt=all_moves_female)
ate_treatment_btwnmalefemale(dt=all_moves_female)

kable(complyratesbytreatment(dt=all_moves_female))
```

Conduct t-tests by move

```{r two-sample t-test for male subjects only}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_male_tm <- t_test_result(all_moves_male, 'TM', 'C')
t_test_result_male_tf <- t_test_result(all_moves_male, 'TF', 'C')
t_test_result_male_t <- t_test_result(all_moves_male, 'TM', 'TF')

t_test_result_male_tm
t_test_result_male_tf
t_test_result_male_t
```



```{r two-sample t-test for female subjects only}
# conduct a two-sample t-test and save results to t_test_result variable
t_test_result_female_tm <- t_test_result(all_moves_female, 'TM', 'C')
t_test_result_female_tf <- t_test_result(all_moves_female, 'TF', 'C')
t_test_result_female_t <- t_test_result(all_moves_female, 'TM', 'TF')

t_test_result_female_tm
t_test_result_female_tf
t_test_result_female_t
```


Show results of linear regressions run on aggregated data for by-gendered datasets with robust and clustered standard errors.

```{r}
# report the formatted regression results
reg_male <- all_moves_male_agg[,lm(comply_rate ~ as.factor(assignment_status)+age)]
reg_female <- all_moves_female_agg[,lm(comply_rate ~ as.factor(assignment_status)+age)]

stargazer(reg_male, reg_male,reg_female,reg_female,
          type = 'text', 
          se = list(robust_se(reg_male),cluster_se(reg_male,all_moves_male_agg),robust_se(reg_female),cluster_se(reg_female,all_moves_female_agg)),
          add.lines = list(c('Subject Gender','Male','Male','Female','Female'),c('SE', 'Robust', 'Clustered','Robust','Clustered')),
          #omit.stat = c('ser', 'F'),
          header=F)


```
